-----
title: "PA1_template"
author: "Ajit"
date: "July 28, 2016"
output: html_document
------
## Loading and preprocessing the data
#### Read the files from the working directory
```{r echo=TRUE}
rm(list=ls(all=TRUE))
setwd("~/Data")
activity <- read.csv("activity.csv",header = TRUE,sep = ",")
summary(activity)
str(activity)
require(dplyr)


activityRefined<-activity
activityRefined$date <- as.Date(activityRefined$date)
str(activityRefined)

 
activitySummaryByDate <- summarise(group_by(activityRefined,date),StepsDay = sum(steps))


barplot(height=activitySummaryByDate$StepsDay, names.arg=activitySummaryByDate$date,
            xlab="Date", ylab=expression('Total # of Steps'),
            main=expression('Summary by steps'),col=c("red", "green", "blue", "purple"))


library(ggplot2)
```
## What is mean total number of steps taken per day?

#### the total number of steps taken per day
```{r echo=TRUE}
total.steps <- tapply(activityRefined$steps, activityRefined$date, FUN=sum, na.rm=TRUE)

## a histogram of the total number of steps taken each day
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")

## the mean and median of the total number of steps taken per day
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```

## What is the average daily activity pattern?

####  the average number of steps taken, averaged across all days
```{r echo=TRUE}
interval<- activityRefined%>%
        group_by(interval)%>%
        filter(!is.na(steps))%>%
        summarise(avg_steps = mean(steps, na.rm=TRUE))
interval

averages <- aggregate(x=list(steps=activityRefined$steps), by=list(interval=activityRefined$interval),
                      FUN=mean, na.rm=TRUE)
```
#### time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and all days (y-axis)

```{r echo=TRUE}
ggplot(data=averages, aes(x=interval, y=steps)) + geom_line() + xlab("5-minute interval") + ylab("average number of steps taken")
```
#### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
interval[which.max(interval$avg_steps),]
```

## Imputing missing values
#### 1. Calculate total number of missing values in the dataset:
```{r echo=TRUE}
sum(is.na(activityRefined$steps))

## 2. Imputing missing values using mean for each day and 

activityRefined2<- activityRefined
nas<- is.na(activityRefined2$steps)
avg_interval<- tapply(activityRefined2$steps, activityRefined2$interval, mean, na.rm=TRUE, simplify = TRUE)

## 3. Create a new dataset that is equal to the original dataset but with the missing data filled in:
activityRefined2$steps[nas] <- avg_interval[as.character(activityRefined2$interval[nas])]
names(activityRefined2)

## 4. Check if no missing value is appearing:
sum(is.na(activityRefined2))

## 5. Reorder columns (for better understanding of the data):
activityRefined2<- activityRefined2[, c("date", "interval", "steps")]
head(activityRefined2)

#Create histogram of the total number of steps taken each day and 
# Calculate mean and median total number of steps taken per day:

Total_Steps2<- activityRefined2%>%
        group_by(date)%>%
        summarise(total_steps = sum(steps, na.rm=TRUE))
Total_Steps2
## Plot the graph
ggplot(Total_Steps2, aes(x = total_steps)) +
        geom_histogram(fill = "blue", binwidth = 1000) +
        labs(title = "Daily Steps including Missing values", x = "Interval", y = "No. of Steps")

## 4.2 Do these values differ from the estimates from the first part of the assignment? 

Mean_Steps2<- mean(Total_Steps2$total_steps, na.rm=TRUE)
Mean_Steps2
```

#### Question: What is the impact of imputing missing data on the estimates of the total daily number of steps?
##### Answer: The impact of imputing missing data with the average number of steps in the same 5-min interval is that both the mean and the median are same : 10766.19

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}

head(activityRefined2)

## 5.1 Create new varibale called WeekType for Weekday & Weekend:
activityRefined2<- activityRefined2%>%
        mutate(weektype= ifelse(weekdays(activityRefined2$date)=="Saturday" | weekdays(activityRefined2$date)=="Sunday", "Weekend", "Weekday"))

head(activityRefined2)

## Plotting:
interval2<- activityRefined2%>%
        group_by(interval, weektype)%>%
        summarise(avg_steps2 = mean(steps, na.rm=TRUE))
head(interval2)

plot<- ggplot(interval2, aes(x =interval , y=avg_steps2, color=weektype)) +
       geom_line() +
       labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") +
       facet_wrap(~weektype, ncol = 1, nrow=2)
print(plot)
```

####Ans: During the weekday, the test object is more active earlier in the day but the object is more active 
####  throughout the weekends probably because the oject is working during the weekdays, hence moving less during the day.



